\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-03-22 paper.tex
Section on robot control

\subsection{Modeling the knowledge related to the interaction}

We split the interaction situations stemming from the situation assessment and
communication components in two categories: \emph{desires} and
\emph{experiences}.

Experiences comprise of emotions and questions

\subsubsection{Event-driven control}

@@

In order to devise how a given goal can be accomplished, the robot has to
elaborate a plan, \ie a set of actions to be achieved by the robot and its
human partner(s).  We use in our architecture the HATP planner (for Human Aware
Task Planner~\cite{Alili2009}).  HATP is based on a Hierarchical Task Network
(HTN) refinement, which performs an iterative task decomposition into sub-tasks
until reaching atomic actions.  The \emph{planning domain} defines the set of
methods describing how to decompose a task and can be seen as the procedural
knowledge of the robot.

>>>>

\subsection{Desires and experiences}

Our robot execution controllers (OpenPRS-based {\sc Shary} or Python-based {\sc
pyRobots}) have deep integration with the knowledge base. It serves as the primary source of information for semantic-aware decision-making.

We split the interaction situations stemming from the situation assessment and
communication components in two categories: \emph{desires} (performative act)
and \emph{experiences} (assertive act).

\emph{Desires} are typically human orders (``Give me that book''). The nature
of the desired action (to pick, give, look, bring, show...), along with the
action parametrization (what is acted on? who should perform the action? etc.)
are extracted from the knowledge base, and either passed to a task planner
(presented in the next section) or executed if the procedure is directly
available.

\emph{Experiences}, on the other hand, comprise of emotions, states and
questions (when asking a question, we consider the human to be in an
\emph{interrogative state}). When the knowledge base recognizes that an agent
\emph{experiences} a particular emotion or state, the execution controller may
decide to handle it, typically by trying to answer the question or using the
emotional or physical state as a parameter for subsequent actions. As an example,
when the speaker says ``I feel tired'', we change the motion planner
parametrization to lower the effort the human need to provide for the following
joint manipulation tasks. Note that this example has been implemented as a
proof-of-concept. We have not yet tried to define a theoretical framework that
would support action alteration based on the user's experienced states.

\subsection{Event-driven control}

The {\sc Oro} server proposes two paradigms to access its content: RPC-style
queries (based on the standard SPARQL language) or events. In its simplest
form, a module can register an event by passing through an event pattern, which
is basically a partial statement like \stmt{? type Elephant}, and a callback.
Every time a new instance of elephant appears in the knowledge base, the
callback is triggered.

This allows us to write reactive robot controllers with a high level of
expressiveness: for instance, by subscribing to the event \stmt{human1 desires
?action, ?action type Look, ?action hasGoal myself}, we could trigger a
behaviour when the human expresses (through dialogue, gestures...) that he
wants to look at the robot itself.

The robot controller designer does not need to directly care about how this
\emph{desire} is produced (this is delegated to perception modules), it can
focuses on the semantic of the desire.

Note also that we take advantage of the reasoning capabilities of the system:
for example, the goal of the action (\stmt{action hasGoal myself}) may well not
have been explicitly asserted, but inferred by the reasoner based on other
assertions.

@@

Our execution controllers rely on symbolic task planning to convert faraway
desires into a succession of atomic actions. We use in our architecture the
HATP planner (for Human Aware Task Planner~\cite{Alili2009}).  HATP is based on
a Hierarchical Task Network (HTN) refinement, which performs an iterative task
decomposition into sub-tasks until reaching atomic actions.  The \emph{planning
domain} defines the set of methods describing how to decompose a task and can
be seen as the procedural knowledge of the robot.
